Running `Earth_obs_cleaning` on MCD19A2: Results from CONUS

### 2022-04-13 R/CONUS_AOD.Rmd
update CONUS results report

```{r basics, echo=FALSE}
library(targets)
library(data.table)
library(DT)
library(patchwork)
library(ggplot2)
library(sf)
```

@@

### Performance by year
```{r strata_year, echo=FALSE}
dtwrapper(dt[, performance_aod(.SD), keyby = year(aer_date)])
```

@@

## Performance by NOAA climate regions

@@

```{r get_noaa_regions, include=FALSE}
# assign AERONET stations to states
tar_load(aer_stations)
aer_stations_sf <- st_as_sf(aer_stations, coords = c("lon", "lat"), crs = "epsg:4326")
sf::sf_use_s2(FALSE) # we were getting errors about the USA geometry otherwise
aer_stations_sf <- st_join(aer_stations_sf, usa)
dt <- merge(dt, aer_stations_sf, by = "Site_Name")
# some Site_Names don't fall within land polygons for whatever reason
dt[is.na(ID), table(Site_Name)]
# where are the sites that didn't fall within a state?
site_nostate <- st_as_sf(dt[is.na(ID), ][unique(Site_Name), .(Site_Name, `Site_Elevation(m)`, geometry), mult = "first"], crs = "epsg:4326")
ggplot() +
  geom_sf(data = usa, fill = NA) +
  geom_sf(data = site_nostate, color = "red")
# what proportion of data are not in a state?
dt[, proportions(table(is.na(ID)))]
# assign these to the closest state
matched <- st_join(site_nostate,
                   usa,
                   join=nngeo::st_nn, k=1)
# plot these matchups
ggplot() +
  geom_sf(data = usa, fill = NA) +
  geom_sf_text(data = matched, aes(label = toupper(substr(ID, 1, 1))), color = "red") + theme_minimal()

@@

```{r agreement_plot, echo=FALSE}
p1 <- ggplot(dt) +
  aes(AOD_470nm, MCD19_AOD_470nm) +
  geom_abline(linetype = "dashed") +
  geom_point(alpha = 0.01) +
  geom_smooth() +
  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1)) +
  scale_y_continuous("MAIAC AOD 470nm", limits = c(0, 1)) +
  theme_bw() +
  theme(aspect.ratio = 1)

p2  <- ggplot(dt) +
  aes(AOD_470nm, aodhat) +
  geom_abline(linetype = "dashed") +
  geom_point(alpha = 0.01) +
  geom_smooth() +
  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1)) +
  scale_y_continuous("Corrected AOD 470nm", limits = c(0, 1)) +
  theme_bw() +
  theme(aspect.ratio = 1)

@@

>>>>

```{r basics, echo=FALSE}
library(targets)
library(data.table)
library(DT)
library(patchwork)
library(ggplot2)
library(sf)
library(knitr)
```

@@

### Performance by year
```{r strata_year, echo=FALSE, eval=TRUE}
dtwrapper(dt[, performance_aod(.SD), keyby = year(aer_date)])
```

@@

### Performance by NOAA climate regions

@@

```{r get_noaa_regions, include=FALSE}
# assign AERONET stations to states
tar_load(aer_stations)
aer_stations_sf <- st_as_sf(aer_stations, coords = c("lon", "lat"), crs = "epsg:4326")
sf::sf_use_s2(FALSE) # we were getting errors about the USA geometry otherwise
aer_stations_sf <- st_join(aer_stations_sf, usa)
dt <- merge(dt, aer_stations_sf, by = "Site_Name")
# some Site_Names don't fall within land polygons for whatever reason
dt[is.na(ID), table(Site_Name)]
# where are the sites that didn't fall within a state?
site_nostate <- st_as_sf(dt[is.na(ID), ][unique(Site_Name), .(Site_Name, `Site_Elevation(m)`, geometry), mult = "first"], crs = "epsg:4326")
# ggplot() +
#   geom_sf(data = usa, fill = NA) +
#   geom_sf(data = site_nostate, color = "red")
# what proportion of data are not in a state?
dt[, proportions(table(is.na(ID)))]
# assign these to the closest state
matched <- st_join(site_nostate,
                   usa,
                   join=nngeo::st_nn, k=1)
# plot these matchups
# ggplot() +
#   geom_sf(data = usa, fill = NA) +
#   geom_sf_text(data = matched, aes(label = toupper(substr(ID, 1, 1))), color = "red") + theme_minimal()

@@

```{r agreement_plot, warning=FALSE, echo=FALSE}
p1 <- ggplot(dt) +
  aes(AOD_470nm, MCD19_AOD_470nm) +
  geom_abline(linetype = "dashed") +
  geom_point(alpha = 0.01) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1)) +
  scale_y_continuous("MAIAC AOD 470nm", limits = c(0, 1)) +
  theme_bw() +
  theme(aspect.ratio = 1)

p2  <- ggplot(dt) +
  aes(AOD_470nm, aodhat) +
  geom_abline(linetype = "dashed") +
  geom_point(alpha = 0.01) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1)) +
  scale_y_continuous("Corrected AOD 470nm", limits = c(0, 1)) +
  theme_bw() +
  theme(aspect.ratio = 1)

@@

## SHAP interpretation ####
In decreasing order (from Terra collocated cross-validation predictions)
```{r shap_rank, echo=FALSE}
# what do the SHAPs tell us?
tar_load(initial_cv_l2_terra_conus)
shaps <- initial_cv_l2_terra_conus$"shap_score"
# sort by importance
kable(sort(sapply(abs(shaps), mean), decreasing = TRUE), 
      format = "html",
      table.attr = "style='width:40%;'",
      digits = 4, col.names = "Mean Abs SHAP")
